---
title: "Cardio Vascular Disease Project"
author: "Team 1: Afif Mahmassani, Turner Luo, Lijia Ren, Steven Tian"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
library(ezids)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```
# Introudction

# Description of Dataset  
```{r}
cardio_vascular_disease <- read.csv2("cardio_train.csv", header = TRUE)
str(cardio_vascular_disease)
```
Noticed some of the variables should be factor.    
```{r}
df <- cardio_vascular_disease
df$gender <- factor(cardio_vascular_disease$gender)
df$smoke <- factor(cardio_vascular_disease$smoke)
df$alco <- factor(cardio_vascular_disease$alco)
df$active <- factor(cardio_vascular_disease$active)
df$cardio <- factor(cardio_vascular_disease$cardio)
df$cholesterol <- factor(cardio_vascular_disease$cholesterol)
df$weight <- as.integer(cardio_vascular_disease$weight)
df <- na.omit(df)
str(df)
``` 
The variables in the dataset are:   

* `id`: Patient ID   
* `age`: Age (in days)
* `gender`: Gender, 1: Female, 2: Male
* `height`: Height (in cm)
* `weight`: Weight (in kg)
* `ap_hi`: Systolic blood pressure
* `ap_lo`: Diastolic blood pressure
* `cholesterol`: Cholesterol, 1: Normal, 2: Above Normal, 3: Well Above Normal
* `glue`: Glucose, 1: Normal, 2: Above Normal, 3: Well Above Normal
* `smoke`: Whether patient smokes or not, 1: Do not smoke, 2: Smoke
* `alco`: Alcohol intake, 1: Do not drink alcohol, 2: Drink alcohol
* `active`: Physical activity, 1: Do not exercise, 2: Exercise
* `df`: Presence or absence of cardiovascular disease, 1: Absence, 2: Presence


```{r}
xkablesummary(df[,2:13], title = "Summary of Cardio Vascular Disease Data")
```
# Categorical Variables EDA   
# Continuous Variables EDA     
# EDA on the joint effect of influencing factors     
# Model Comparison      
## Logistic Regression   
```{r}
#need correlation plot to select variables
df_num <- df

for(i in 2:13){
  if (!(i %in% c(2, 4:7, 9))){
    df_num[,i] = as.numeric(df_num[,i])
  }
}
cor <- cor(subset(df_num, select = -c(id)), method = "spearman")
loadPkg("corrplot")
corrplot(cor, type="lower", addCoef.col="black", number.cex=0.5, tl.cex=0.7,title="Cardio Vascular Disease Correlation", mar=c(0,0,1,0))
```
  The first model we are performing is logistic regression model. There are 13 variables, first we exclude `id` which is not related to the cardio vascular disease. Then to prevent the multicollinearity, we choose one between `ap_hi` and `ap_lo`. So, we start with a logistic model with 10 variables. We set alpha equal to 0.05, so we should exclude all the variable that have p-value larger than 0.05.   
```{r}
logistic_all <- glm(cardio ~ age + gender + height + weight + ap_hi + ap_lo + cholesterol + gluc + smoke + alco + active, data = df, family = "binomial")
summary(logistic_all)
```
With p-value of 0.49, we delete `gender`.  
```{r}
logistic_1 <- glm(cardio ~ age + height + weight + ap_hi + ap_lo + cholesterol + gluc + smoke + alco + active, data = df, family = "binomial")
summary(logistic_1)
```
```{r}
coef1 <- exp(coef(logistic_1))
xkabledply(as.table(coef1), title = "Exponential of coefficiant in logistic model")
```

AUC and ROC:   
```{r}
loadPkg("pROC")
prob <- predict(logistic_1, type = "response")
df$prob <- prob
h <- roc(cardio ~ prob, data = df)
auc(h)
plot(h)
```
