---
title: "Cardio Vascular Disease Project"
output: html_document
date: '2022-08-06'
---

```{r}
library(ezids)
library(ggplot2)
library(dplyr)
```

```{r}
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```

```{r}
CVD <- data.frame(read.csv("cardio_train.csv", sep=";"))
View(CVD)
```


## Chapter 1: Introudction to Cardiovascular Diseases (CVDs)

# 1.1 Basic Information About Cardiovascular Diseases (CVDs)

Cardiovascular Disease is a general name for numerous types of diseases including heart attack, stroke, and heart failure. In 2019 an estimated 17.9 million people died from Cardiovascular Diseases and 85% were due to heart attack and stroke.

# 1.2 Reasons of Choosing Cardiovascular Disease.
As living standards are improving over the past few decades, having a healthy lifestyle and avoiding serious illnesses have become people’s priorities and Cardiovascular diseases are the leading cause of death globally our group is interested in finding out some facts about Cardiovascular Disease such as what exactly Cardiovascular Diseases are. What are some factors that are correlate to Cardiovascular Diseases, among those factors what have the strongest correlation to Cardiovascular Diseases? And would a healthy lifestyle help to prevent people from having cardiovascular disease? Therefore, our group choose this dataset about Cardiovascular Diseases

## Chapter 2: Summary of the Dataset

# 2.1:About the Dataset
The Cardiovascular Diseases dataset is extracted from Kaggle. It contains information from 70,000 records of patients that provided their age, height, weight, physical condition, and habits. Studying such data can help the public and health organizations identify potential factors that could lead to a higher chance of getting Cardiovascular Diseases, and develop appropriate strategies to reduce the number of patients.
The dataset is cardio_train.csv. Before doing any analysis on this dataset, we did some research about what every single variable in the dataset represents, why they are included in this dataset, and what variables we will use in our model. With all those information government agents and the public will know how to adjust their habits or lifestyle to prevent getting Cardiovascular Diseases. Therefore, it is crucial to figure out current problems, and then publish the information.

# 2.2: Variables
```{r}
str(CVD)
```

1.`id`: Respondents'id | id | int

2.`Age`: Respondents' age | age | int (days)

3.`Height`: Respondents' height | height | int (cm)

4.`Weight`: Respondents' weight | weight | num (kg)

5.`Gender`: Respondents'gender | gender  | int (1:Feale|2:Male)

6.`Systolic blood pressure`: Respondents' systolic blood pressure | ap_hi | int 

7.`Diastolic blood pressure`: Respondents' diastolic blood pressure | ap_lo | int

8.`Cholesterol`: Respondents' cholesterol level | cholesterol | int (1: normal, 2: above normal, 3: well above normal)

9.`Glucose`: Respondents' glucose level | gluc | int (1: normal, 2: above normal, 3: well above normal)

10.`Smoking`: Whether respondent smokes or not | smoke | int (0:Do not smoke, 1:Smoke) 

11.`Alcohol intake`: Whether respondent drink or not | alco | int (0:Do not drink, 1: Drink)

12.`Physical activity`: Respondents' do not physical activity regularly | active | int (0:Workout regularly, 1: No workout regularly)

13.`Presence or absence of cardiovascular disease`: Respondent present cardiovascular disease | cardio | int (0:No CVD currently present, 1:CVD currently present)

## Chapter 3: Descriptive Statistics

# 3.1 Categorical Variables EDA
# 3.1.1 Are there any common in CVD patients?
We choose several factors from the dataset and check if there are some factors that have a significant difference between CVD presented and CVD not presented.All the tables below will use percentage. 

We start with smoke history and whether CVD are currently presented or not.
Smoke
```{r,results='markup'}
library(dplyr)
smoke_con<-aggregate(CVD$cardio, by=list(CVD$smoke,CVD$cardio), FUN=length)
smoke_con<-rename(smoke_con,Smoke = Group.1,CVD_presence = Group.2,CVD = x)
smoke_con$Smoke[smoke_con$Smoke=="0"] <- 'Not smoke'
smoke_con$Smoke[smoke_con$Smoke=="1"] <- 'Smoke'
smoke_con$CVD_presence[smoke_con$CVD_presence=="0"] <- 'Not Present'
smoke_con$CVD_presence[smoke_con$CVD_presence=="1"] <- 'Present'
xkabledply(smoke_con)
```

alcohol intake 
```{r,results='markup'}
alco_con<-aggregate(CVD$cardio,by=list(CVD$alco,CVD$cardio), FUN=length)
alco_con<-rename(alco_con,alco = Group.1,CVD_presence = Group.2,CVD = x)
alco_con$alco[alco_con$alco=="0"] <- 'Not drink'
alco_con$alco[alco_con$alco=="1"] <- 'Drink'
alco_con$CVD_presence[alco_con$CVD_presence=="0"] <- 'Not Present'
alco_con$CVD_presence[alco_con$CVD_presence=="1"] <- 'Present'
xkabledply(alco_con)
```

Physical Activity
```{r,results='markup'}
active_con<-aggregate(CVD$cardio, by=list(CVD$active,CVD$cardio), FUN=length)
active_con<-rename(active_con,active = Group.1,CVD_presence = Group.2,CVD = x)
active_con$active[active_con$active=="0"] <- 'Physical Activity'
active_con$active[active_con$active=="1"] <- 'Not Physical Activity'
active_con$CVD_presence[active_con$CVD_presence=="0"] <- 'Not Present'
active_con$CVD_presence[active_con$CVD_presence=="1"] <- 'Present'
xkabledply(active_con)
```

# 3.1.2 Plot comparing different factors

Plot of Smoke & CVD
```{r,results='markup'}
smoke_pic <- data.frame(table(CVD$smoke,CVD$cardio))
names(smoke_pic) <- c("Smoke","Cardio","Count")
ggplot(data=smoke_pic, aes(x=Smoke, y=Count/sum(Count)*100, fill=Cardio)) + geom_bar(stat="identity")+scale_x_discrete(labels=c("0" = "Not smoke", "1" = "Smoke"))+scale_fill_discrete(labels=c("0" = "No cvd presented", "1" = "CVD presented"))+ggtitle("Percentage of Smoke history and present of CVDs") + ylab("Percentage")
```

Plot of alcohol intake & CVD
```{r,results='markup'}
alco_pic <- data.frame(table(CVD$alco,CVD$cardio))
names(alco_pic) <- c("alco","Cardio","Count")
ggplot(data=alco_pic, aes(x=alco, y=Count/sum(Count)*100, fill=Cardio)) + geom_bar(stat="identity")+scale_x_discrete(labels=c("0" = "Not Drink", "1" = "Drink"))+scale_fill_discrete(labels=c("0" = "No cvd presented", "1" = "CVD presented"))+ggtitle("Percentage of alcohol intake and present of CVDs") + ylab("Percentage")
```

Plot of physical activity & CVD
```{r,results='markup'}
active_pic <- data.frame(table(CVD$active,CVD$cardio))
names(active_pic) <- c("active","Cardio","Count")
ggplot(data=active_pic, aes(x=active, y=Count/sum(Count)*100, fill=Cardio)) + geom_bar(stat="identity")+scale_x_discrete(labels=c("0" = "Workout regularly", "1" = "Not workout regularly"))+scale_fill_discrete(labels=c("0" = "No cvd presented", "1" = "CVD presented"))+ggtitle("Percentage of workout and present of CVDs") + ylab("Percentage")
```

# 3.1.3 Chi-Square Tests

Since there are 

Chi-square test for Smoke & CVD presence
```{r,results='hide'}
smoke_cvd = table(CVD$cardio, CVD$smoke)
```

```{r,results='markup'}
sctest<- chisq.test(smoke_cvd)
sctest
```

Chi-square test for Alcohol intake & CVD presence
```{r,results='hide'}
al_cvd = table(CVD$cardio, CVD$alco)
```
```{r,results='markup'}
alctest<- chisq.test(al_cvd)
alctest
```

Chi-square test for Physical Activity & CVD presence
```{r,results='hide'}
act_cvd = table(CVD$cardio, CVD$active)
```
```{r,results='markup'}
act_test<- chisq.test(act_cvd)
act_test
```



# Chapter 4: Continuous Variables EDA to Cardiovascular Diseases (CVDs)

After a brief overview of the dataset, it tells that there are 3 continuous variables, age, height and weight.

## 4.1 KDE plot of continuous variables



KDE plot represents the data using a continuous probability density curve in one or more dimensions. We can easily observe the distribution of samples with kde plot

```{r init, include=F}
library(ezids)
library(ggplot2)
library(dplyr)
library(ggpubr)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```



```{r setup, include=FALSE}

CVD <- data.frame(read.csv("cardio_train.csv", sep=";"))
View(CVD)


```


```{r setup, include=FALSE}

str(CVD)

```


```{r setup, include=FALSE}


CVD$cardio <- factor(CVD$cardio)
str(CVD)

```






```{r setup, include=FALSE}

CVD_new<- data.frame(CVD)
CVD_new
```

```{r setup, include=FALSE}


levels(CVD_new$cardio) <- c("No CVD Present", "CVD Present") 
str(CVD_new)

```


```{r}



CVD_new$cardio[CVD_new$cardio=="0"] <- 'No CVD Present'
CVD_new$cardio[CVD_new$cardio=="1"] <- 'CVD Present'
CVD_new
```







```{r, results="markup"}
cols<-c("No CVD Present"="green","CVD Present"="blue")

```







```{r, results="markup"}
age_kdeplot <- ggplot(data = CVD_new, aes(x = age, color =cardio)) + 
            geom_density(aes(fill = cardio), alpha = 0.8) + 
             scale_fill_manual(values =cols) +
              labs(title="KDEplot for age") +
               labs(x="age", y="density") +
                theme(legend.position="top")
age_kdeplot


```


```{r, results="markup"}
age_kdeplot <- ggplot(data = CVD_new, aes(x = height, color = cardio)) + 
            geom_density(aes(fill = cardio), alpha = 0.8) + 
             scale_fill_manual(values =cols) +
              labs(title="KDEplot for height") +
               labs(x="height", y="density") +
                theme(legend.position="top")
age_kdeplot


```



```{r, results="markup"}
age_kdeplot <- ggplot(data = CVD_new, aes(x = weight, color = cardio)) + 
            geom_density(aes(fill = cardio), alpha = 0.8) + 
             scale_fill_manual(values= cols) +
              labs(title="KDEplot for weight") +
               labs(x="weight", y="density") +
                theme(legend.position="top")
age_kdeplot


```

As we can see from kdeplot for age, people with higher age are more likely to have CVD. And from kdeplot for height  and weight ,  we can find that  people with CVD and without CVD have very similar distributions. From these 3 kde plots, which means,CVD may be positive correlated with age. Finally, weight may only make a little attribution to people presenting CVD.

## 4.2 Logistic regression

The logit model is often used for classification and predictive analytics. Logistic regression estimates the probability of an event occurring, such as voted or didn’t vote, based on a given dataset of independent variables. Since the outcome is a probability, the dependent variable is bounded between 0 and 1(binary outcome). In this case, "cardio"  is the binary outcome; as s result, we choose logistic regression to predict.


```{r, results="markup"}


lm1 <- glm(cardio~age, family = binomial(link = "logit"), data = CVD)
lm2 <- glm(cardio~age + weight, family = binomial(link = "logit"), data = CVD)
lm3 <- glm(cardio~age+ height + weight, family = binomial(link = "logit"), data = CVD)
anovat <- anova(lm1,lm2,lm3, test="LRT")
anovat

```

As it is shown above, model 2 and 3 is better than model 1 . In addition, model 2 and 3 is at 100% significant level.



## 4.3 AUC and ROC Curve

We can use AUC and ROC to measure model 2 and model 3. AUC (Area Under The Curve) - ROC (Receiver Operating Characteristics) curve is a performance measurement for the classification problems at various threshold settings. ROC is a probability curve and AUC represents the degree or measure of separability. It tells how much the model is capable of distinguishing between classes. Higher the AUC, the better the model is at predicting 0 classes as 0 and 1 classes as 1. For summarize, the Higher the AUC, the better the model is.

For model 2:


```{r, results="markup"}
prob <- predict(lm2,CVD, type = c("response"))
CVD$prob <- prob
library(pROC)
g<- roc(cardio~prob, data = CVD)
plot(g, main = "ROC curve of model 2")
auc(CVD$cardio, prob)
```
The AUC value is 0.668. As 0.668 is much less than 1, this model is not good enough.


Below is the summary of model 2.

````{r, results="markup"}
xkabledply(lm2, title="summary of model 2")
```

For model 3:



```{r, results="markup"}
prob1 <- predict(lm3,CVD, type = c("response"))
CVD$prob1 <- prob1
library(pROC)
g<- roc(cardio~prob1, data = CVD)
plot(g, main = "ROC curve of model 3")
auc(CVD$cardio, prob1)
```
The AUC value is 0.671. As 0.671 is much less than 1 and 0.671 is more than 0.668, this model is not good enough but is better than model 2.


Below is the summary of model 3.

```{r, results="markup"}
xkabledply(lm3, title="summary of model 3")
```

In the 3 continuous variables, `cardio` has positive correlation with `age` and `weight` ; it also has negative correlation with`height` . It means when a person with  higher `weight` and shorter `height`, he has more probabilities to have Cardio Vascular Disease. And `age` is not significant influence factor to people `cardio` rates in 3 continuous variables.


# 5 Chapter 5: EDA on the joint effect of influencing factors


From the results of above variables, we could find that some of the above variables have a significant effect on CVD. Next, we choose to graph the groups of variables with relatively large correlation coefficients one by one to explore how they affect the Cardio Vascular Disease.

## 5.1 Simple correlations

As most of variables are factors, we check their Spearman correlations.

```{r init, include=F}
show(CVD_new)
```




```{r, results="markup"}

CVDNum = CVD

str(CVDNum)
```

```{r, results="markup"}

# convert categorical variable as numeric for spearman method
for(i in 1:13){
  # age, height, weight
  if (!(i %in% c(2, 4, 5))){
    CVDNum[,i] = as.numeric(CVDNum[,i])
              }
}
str(CVDNum)

```

```{r, results="markup"}

# corrplot with spearman method for categorical variables
CVDcor <- cor(subset(CVDNum, select=-c(id, prob)), method="spearman")
#CVDcor
loadPkg("corrplot")
#corrplot.mixed(CVDcor, tl.pos = "lt", number.cex = .5, tl.cex=0.8)
corrplot(CVDcor, type="lower", addCoef.col="black", number.cex=0.5, tl.cex=0.7,title="Spearman Correlation for CVD", mar=c(0,0,1,0))

```
Larger circle means higher correlation. We can see that CVD has negative correlation with cholesterol and height, which means that person who has shorter height or has high cholesterol is less likely to have Cardio Vascular Disease. Person who smoke and has high level cholesterol is also more likely to have CVD. So it makes sense that cholesterol  and smoke have positive correlation, which means most person who smoke also have high level cholesterol .

## 5.2 Will Age influence the CVD rate with other variables?


```{r init, include=F}
CVD_new$gluc<- factor(CVD_new$gluc)
CVD_new$gender<- factor(CVD_new$gender)
CVD_new$smoke<- factor(CVD_new$smoke)
CVD_new$alco<- factor(CVD_new$alco)
CVD_new$active <- factor(CVD_new$active)
str(CVD_new)

```

```{r init, include=F}
CVD_factor<- data.frame(CVD_new)
CVD_factor
```

```{r init, include=F}

library(patchwork)
```

```{r, results="markup"}


a1 <- ggplot(CVD_factor,aes(x = smoke , y = age , fill = cardio)) +
  geom_violin(alpha = 0.5, aes(linetype=NA)) +
  xlab("Smoke") + ylab("Age")


a2 <- ggplot(CVD_factor,aes(x = alco , y = age , fill = cardio)) +
  geom_violin(alpha = 0.5, aes(linetype=NA)) +
  xlab("Alco") + ylab("Age")


a3 <- ggplot(CVD_factor, aes(x=height, y=age, color= cardio)) +
  geom_point(size=.5,alpha=0.4)


```

```{r, results="markup"}


a3/ (a1+a2) + plot_annotation(title = 'Age Plot')
```
First, Age and height show no correlation.

Second, for Smoke, the sample groups of smoke("1") and order age have higher CVD rates .

Finally, for cholesterol, people are also prone to CVD when drink alco("1") and order age.

## 5.3 Will Height influence the CVD rate with other variables?


```{r, results="markup"}


h1 <- ggplot(CVD_factor,aes(x = active , y = height , fill = cardio)) +
  geom_violin(alpha = 0.5, aes(linetype=NA)) +
  xlab("Active") + ylab("Height")



h2 <- ggplot(CVD_factor, aes(x=weight, y=height, color= cardio)) +
  geom_point(size=0.5,alpha=0.4)


```



```{r, results="markup"}


h1/ h2 + plot_annotation(title = 'Height Plot')
```

First, for Active, the sample groups of active("1") have little higher CVD rates.

Second, weight and height show a little positive correlation .



## 5.4 Will Weight influence the CVD rate with other variables?


```{r, results="markup"}


w1 <- ggplot(CVD_factor,aes(x = active , y = weight , fill = cardio)) +
  geom_violin(alpha = 0.5, aes(linetype=NA)) +
  xlab("Active") + ylab("Weight")


w2 <- ggplot(CVD_factor,aes(x = alco , y = weight , fill = cardio)) +
  geom_violin(alpha = 0.5, aes(linetype=NA)) +
  xlab("Alco") + ylab("Weight")




```



```{r, results="markup"}


w1/ w2 + plot_annotation(title = 'Weight Plot')
```

First, for Active, the sample groups of  active("1") have little higher CVD rates 

Second, for Alco, the sample groups of no alco ("0") have little higher CVD rates.

#Models comparison: Logistic Regression

```{r}
df <- cardio_vascular_disease
df$id <- NULL
df$gender <- factor(cardio_vascular_disease$gender)
df$smoke <- factor(cardio_vascular_disease$smoke)
df$alco <- factor(cardio_vascular_disease$alco)
df$active <- factor(cardio_vascular_disease$active)
df$cardio <- factor(cardio_vascular_disease$cardio)
df$cholesterol <- factor(cardio_vascular_disease$cholesterol)
df$gluc <- factor(cardio_vascular_disease$gluc)
df$weight <- as.integer(cardio_vascular_disease$weight)

levels(df$cardio) <- c("No", "Yes")
levels(df$gender) <- c("Female", "Male")
levels(df$cholesterol) <- c("Normal", "Above Normal", "Well Above Normal")
levels(df$gluc) <- c("Normal", "Above Normal", "Well Above Normal")
levels(df$smoke) <- c("No", "Yes")
levels(df$alco) <- c("No", "Yes")
levels(df$active) <- c("No", "Yes")
df <- na.omit(df)

str(df)
``` 

```{r, results='markup'}
logistic_all <- glm(cardio ~ age + gender + height + weight + ap_hi + ap_lo + cholesterol + gluc + smoke + alco + active, data = df, family = "binomial")
summary(logistic_all)
```

```{r}
logistic_1 <- glm(cardio ~ age + height + weight + ap_hi + ap_lo + cholesterol + gluc + smoke + alco + active, data = df, family = "binomial")
summary(logistic_1)
```

```{r, results='markup'}
logistic_2 <- glm(cardio ~ age + height + weight + ap_hi + ap_lo + cholesterol + smoke + alco + active, data = df, family = "binomial")
summary(logistic_2)
```
```{r}
expcoef = exp(coef(logistic_2))
summary(expcoef)
xkabledply( as.table(expcoef), title = "Exponential of coefficients in Cardio")
```

```{r, results='markup'}
loadPkg("regclass")
confusion_matrix(logistic_2)
cfmatrix1 = confusion_matrix(logistic_2)
accuracy1 <- (cfmatrix1[1,1]+cfmatrix1[2,2])/cfmatrix1[3,3]
precision1 <- cfmatrix1[2,2]/(cfmatrix1[2,2]+cfmatrix1[1,2])
recall1 <- cfmatrix1[2,2]/(cfmatrix1[2,2]+cfmatrix1[2,1])
specificity1 <- cfmatrix1[1,1]/(cfmatrix1[1,1]+cfmatrix1[1,2])
F1_score1 <- 2*(precision1)*(recall1)/(precision1 + recall1)
accuracy1
precision1
recall1
specificity1
F1_score1
```
```{r}
xkabledply( confusion_matrix(logistic_2), title = "Confusion matrix from Logit Model" )
```

```{r, results='markup'}
loadPkg("pROC")
prob <- predict(logistic_2, type = "response")
df$prob <- prob
h <- roc(cardio ~ prob, data = df)
auc(h)
plot(h)
```
Not a good fit.   
#Feature Selection   
```{r}
loadPkg("leaps")
reg.leaps <- regsubsets(cardio ~ ., data = df, nbest = 1, method = "exhaustive")
plot(reg.leaps, scale = "adjr2", main = "Adjusted R^2")
```
```{r}
plot(reg.leaps, scale = "bic", main = "BIC")
```
```{r}
plot(reg.leaps, scale = "Cp", main = "Cp")
```
Close results.   

# Models Comparison: KNN

```{r}
df <- cardio_vascular_disease
df$id <- NULL
df$gender <- factor(cardio_vascular_disease$gender)
df$smoke <- factor(cardio_vascular_disease$smoke)
df$alco <- factor(cardio_vascular_disease$alco)
df$active <- factor(cardio_vascular_disease$active)
df$cardio <- factor(cardio_vascular_disease$cardio)
df$cholesterol <- factor(cardio_vascular_disease$cholesterol)
df$gluc <- factor(cardio_vascular_disease$gluc)
df$weight <- as.integer(cardio_vascular_disease$weight)

levels(df$cardio) <- c("No", "Yes")
levels(df$gender) <- c("Female", "Male")
levels(df$cholesterol) <- c("Normal", "Above Normal", "Well Above Normal")
levels(df$gluc) <- c("Normal", "Above Normal", "Well Above Normal")
levels(df$smoke) <- c("No", "Yes")
levels(df$alco) <- c("No", "Yes")
levels(df$active) <- c("No", "Yes")
df <- na.omit(df)

str(df)
``` 

```{r}
df_num <- df
for(i in 1:12){
  if (!(i %in% c(1,3:6))){
    df_num[,i] = as.numeric(df_num[,i])
  }
}
str(df_num)
```

```{r}
#convert variables to numeric
scale1 <- subset(df_num, select = -cardio)
scale1$cardio <- df$cardio
str(scale1)
scale1[c(1,3,4,5,6)] <- as.data.frame(scale(scale1[c(1,3,4,5,6)], center = TRUE, scale = TRUE))
set.seed(1)
df_sample <- sample(2, nrow(scale1), replace = TRUE, prob = c(0.75,0.25))
df_train <- scale1[df_sample==1, 1:11]
df_test <- scale1[df_sample==2, 1:11]
df.trainLabels <- df[df_sample==1, 12]
df.testLabels <- df[df_sample==2, 12]
str(df_train)
str(df_test)
```

```{r, echo=FALSE}
loadPkg("class")
chooseK = function(k, train_set, val_set, train_class, val_class){
  
  set.seed(1)
  class_knn = knn(train = train_set,   
                  test = val_set,      
                  cl = train_class,   
                  k = k)                 
  tab = table(class_knn, val_class)
  accu = sum(tab[row(tab) == col(tab)]) / sum(tab)                         
  cbind(k = k, accuracy = accu)
}
```

```{r, results='markup'}
knn_different_k = sapply(seq(1, 21, by = 2),  
                         function(x) chooseK(x, 
                                             train_set = df_train,
                                             val_set = df_test,
                                             train_class = df.trainLabels,
                                             val_class = df.testLabels))
knn_different_k = data.frame(k = knn_different_k[1,],
                             accuracy = knn_different_k[2,])
library("ggplot2")
ggplot(knn_different_k,
       aes(x = k, y = accuracy)) +
  geom_line(color = "orange", size = 1.5) +
  geom_point(size = 3) + 
  labs(title = "accuracy vs k")

xkabledply((knn_different_k))
```
`k=17` is a decent choice.
```{r, results='markup'}
pred <- knn(train = df_train, test = df_test, cl=df.trainLabels, k=17)
pred

loadPkg("gmodels")
churnPredCross <- CrossTable(df.testLabels, pred, prop.chisq = FALSE)
```
accuracy = 0.6648     
precision = 0.6809    
recall = 0.6348     
specificity = 0.6954        
F1_score = 0.657      

### KNN with selected variables    
Remove `gender` and `gluc`.   
```{r}
df_num2 <- subset(df_num, select = -c(gender, gluc, prob))
str(df_num2)
scale2 <- df_num2
scale2[1:5] <- scale(df_num2[1:5], center = TRUE, scale = TRUE)
scale2$cardio <- df$cardio
str(scale2)
```

```{r}
set.seed(1)
df_sample2 <- sample(2, nrow(scale2), replace = TRUE, prob = c(0.75,0.25))
df_train2 <- scale2[df_sample2==1, 1:9]
df_test2 <- scale2[df_sample2==2, 1:9]
df.trainLabel2 <- scale2[df_sample2==1, 10]
df.testLabel2 <- scale2[df_sample2==2, 10]
```

### Select K    
```{r, results='markup'}
knn_different_k2 = sapply(seq(1, 21, by = 2),  
                         function(x) chooseK(x, 
                                             train_set = df_train2,
                                             val_set = df_test2,
                                             train_class = df.trainLabel2,
                                             val_class = df.testLabel2))
str(knn_different_k2)
knn_different_k2 = data.frame(k = knn_different_k2[1,],
                             accuracy = knn_different_k2[2,])
library("ggplot2")
ggplot(knn_different_k2,
       aes(x = k, y = accuracy)) +
  geom_line(color = "orange", size = 1.5) +
  geom_point(size = 3) + 
  labs(title = "accuracy vs k")
xkabledply((knn_different_k2))
```

We should better select value `k=13` here as it has the highest accuracy.  
```{r}
pred2 <- knn(train = df_train2, test = df_test2, cl = df.trainLabel2, k = 13)
knn.roc.prob <- attr(knn(train = df_train2, test = df_test2, cl = df.trainLabel2, k = 13, prob = T), 'prob')
pred2
```

```{r, results='markup'}
churnPredCross2 <- CrossTable(df.testLabel2, pred2, prop.chisq = FALSE)
```

accuracy = 0.6717
precision = 0.6877
recall = 0.6432
specificity = 0.701
F1_score = 0.6616

#Comparison
```{r, results='markup'}
tab <- matrix(c(0.6583, 0.6733, 0.6308, 0.6866, 0.6514, 0.6717, 0.6877, 0.6432, 0.701, 0.6616, 0.722, 0.743, 0.677, 0.767, 0.708), ncol = 5, byrow = TRUE)
colnames(tab) <- c("Accuracy", "Precision", "Recall", "Specificity", "F1 Score")
rownames(tab) <- c("KNN with all variables", "KNN with selected variables", "Logistic with selected variables")
tab <- as.table(tab)
xkabledply(tab, "Models Comparison")
```


# Models comparison: Classification Tree

```{r}
df <- cardio_vascular_disease
df$id <- NULL
df$gender <- factor(cardio_vascular_disease$gender)
df$smoke <- factor(cardio_vascular_disease$smoke)
df$alco <- factor(cardio_vascular_disease$alco)
df$active <- factor(cardio_vascular_disease$active)
df$cardio <- factor(cardio_vascular_disease$cardio)
df$cholesterol <- factor(cardio_vascular_disease$cholesterol)
df$gluc <- factor(cardio_vascular_disease$gluc)
df$weight <- as.integer(cardio_vascular_disease$weight)

levels(df$cardio) <- c("No", "Yes")
levels(df$gender) <- c("Female", "Male")
levels(df$cholesterol) <- c("Normal", "Above Normal", "Well Above Normal")
levels(df$gluc) <- c("Normal", "Above Normal", "Well Above Normal")
levels(df$smoke) <- c("No", "Yes")
levels(df$alco) <- c("No", "Yes")
levels(df$active) <- c("No", "Yes")
df <- na.omit(df)

str(df)
``` 

```{r}
df_num <- df
for(i in 1:12){
  if (!(i %in% c(1,3:6))){
    df_num[,i] = as.numeric(df_num[,i])
  }
}
df_num$cardio <- df$cardio
str(df_num)
```

```{r}
loadPkg("rpart")
loadPkg("caret")



# create an empty dataframe to store the results from confusion matrices
confusionMatrixResultDf = data.frame(Depth = numeric(0), Accuracy  = numeric(0), Sensitivity = numeric(0), Specificity = numeric(0), 
                                      Pos.Pred.Value = numeric(0), Neg.Pred.Value = numeric(0), Precision = numeric(0), Recall = numeric(0),
                                      F1 = numeric(0), Prevalence = numeric(0), Detection.Rate = numeric(0), Detection.Prevalence = numeric(0),
                                      Balanced.Accurary = numeric(0), row.names = NULL)

for (deep in 2:15) {
  kfit <- rpart(cardio ~ age + weight + ap_lo + height + cholesterol, data = df_num, method="class", control = list(maxdepth = deep) )
  # 
  cm = confusionMatrix( predict(kfit, type = "class"), reference = df_num[, "cardio"] ) # from caret library
  # 
  cmaccu = cm$overall['Accuracy']
  # print( paste("Total Accuracy = ", cmaccu ) )
  # 
  cmt = data.frame(Depth=deep, Accuracy = cmaccu, row.names = NULL ) # initialize a row of the metrics 
  cmt = cbind( cmt, data.frame( t(cm$byClass) ) ) # the dataframe of the transpose, with k valued added in front
  confusionMatrixResultDf = rbind(confusionMatrixResultDf, cmt)
  # print("Other metrics : ")
}

unloadPkg("caret")
```
```{r}
xkabledply(confusionMatrixResultDf, title = "Cardio Classificantion Trees summary with varying MaxDepth")
```

```{r}
set.seed(1000)
cardiofit <- rpart(cardio ~ age + weight + ap_lo + height + cholesterol, data = df_num, method = "class", control = list(maxdepth=7))
printcp(cardiofit)
plotcp(cardiofit)
summary(cardiofit)
```

```{r}
plot(cardiofit, uniform = TRUE, main = "Classification Tree for Cardio")
text(cardiofit, use.n = TRUE, all = TRUE, cex = 1)
```
```{r}
loadPkg("caret")
cm = confusionMatrix( predict (cardiofit, type = "class"), reference = df_num[, "cardio"])
print('Overall: ')
cm$overall
```
```{r}
print('Class')
cm$byClass
```
```{r}
xkabledply(cm$table, "confusion matrix")
```
```{r}
loadPkg("rpart.plot")
loadPkg("rattle")
rpart.plot(cardiofit)
fancyRpartPlot(cardiofit)
```
#ROC Curve   

```{r}
library(rpart)
rp <- rpart(cardio ~., data = df_num)
library(ROCR)
pred <- prediction(predict(cardiofit, type = "prob")[,2], df_num$cardio)
tree.predict.prob <- predict(cardiofit, type = "prob")[,2]
plot(performance(pred, "tpr", "fpr"), main = "ROC Cardio")
auc = performance(pred, 'auc')
slot(auc, 'y.values')
```

```{r}
tab1 <- matrix(c(0.672, 0.688, 0.643, 0.701, 0.662, 0.722, 0.743, 0.677, 0.767, 0.708, 0.702, 0.677, 0.771, 0.632, 0.721), ncol = 5, byrow = TRUE)
colnames(tab1) <- c("Accuracy", "Precision", "Recall", "Specificity", "F1 Score")
rownames(tab1) <- c("KNN with selected variables", "Logistic with selected variables", "Classification Tree with selected variables")
tab1 <- as.table(tab1)
xkabledply(tab1, "Model Comparison")
```





