---
title: "Cardio Vascular Disease Project"
author: "Team 1"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=F}
library(ezids)
library(ggplot2)
library(dplyr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
```

```{r}
CVD <- data.frame(read.csv("cardio_train.csv", sep=";"))
View(CVD)
```


## Chapter 1: Introudction to Cardiovascular Diseases (CVDs)

# 1.1 Basic Information About Cardiovascular Diseases (CVDs)

Cardiovascular Disease is a general name for numerous types of diseases including heart attack, stroke, and heart failure. In 2019 an estimated 17.9 million people died from Cardiovascular Diseases and 85% were due to heart attack and stroke.

# 1.2 Reasons of Choosing Cardiovascular Disease.
As living standards are improving over the past few decades, having a healthy lifestyle and avoiding serious illnesses have become people’s priorities and Cardiovascular diseases are the leading cause of death globally our group is interested in finding out some facts about Cardiovascular Disease such as what exactly Cardiovascular Diseases are. What are some factors that are correlate to Cardiovascular Diseases, among those factors what have the strongest correlation to Cardiovascular Diseases? And would a healthy lifestyle help to prevent people from having cardiovascular disease? Therefore, our group choose this dataset about Cardiovascular Diseases

## Chapter 2: Summary of the Dataset

# 2.1:About the Dataset
The Cardiovascular Diseases dataset is extracted from Kaggle. It contains information from 70,000 records of patients that provided their age, height, weight, physical condition, and habits. Studying such data can help the public and health organizations identify potential factors that could lead to a higher chance of getting Cardiovascular Diseases, and develop appropriate strategies to reduce the number of patients.
The dataset is cardio_train.csv. Before doing any analysis on this dataset, we did some research about what every single variable in the dataset represents, why they are included in this dataset, and what variables we will use in our model. With all those information government agents and the public will know how to adjust their habits or lifestyle to prevent getting Cardiovascular Diseases. Therefore, it is crucial to figure out current problems, and then publish the information.

# 2.2: Variables

There are a total of 13 variables in this dataset, including integer, and numeric, we will only use part of the variables in our dataset. There are some outliers in this model. However, overall they do not have a strong impact on our model, and we decided not to remove them. Also for our model, we did some cleanup and convert some of the variables into factor variables so we can better process them.  

```{r}
str(CVD)
```

1.`id`: Respondents'id | id | int

2.`Age`: Respondents' age | age | int (days)

3.`Height`: Respondents' height | height | int (cm)

4.`Weight`: Respondents' weight | weight | num (kg)

5.`Gender`: Respondents'gender | gender  | int (1:Feale|2:Male)

6.`Systolic blood pressure`: Respondents' systolic blood pressure | ap_hi | int 

7.`Diastolic blood pressure`: Respondents' diastolic blood pressure | ap_lo | int

8.`Cholesterol`: Respondents' cholesterol level | cholesterol | int (1: normal, 2: above normal, 3: well above normal)

9.`Glucose`: Respondents' glucose level | gluc | int (1: normal, 2: above normal, 3: well above normal)

10.`Smoking`: Whether respondent smokes or not | smoke | int (0:Do not smoke, 1:Smoke) 

11.`Alcohol intake`: Whether respondent drink or not | alco | int (0:Do not drink, 1: Drink)

12.`Physical activity`: Respondents' do not physical activity regularly | active | int (0:Workout regularly, 1: No workout regularly)

13.`Presence or absence of cardiovascular disease`: Respondent present cardiovascular disease | cardio | int (0:No CVD currently present, 1:CVD currently present)

## Chapter 3: Descriptive Statistics

# 3.1 Categorical Variables EDA
# 3.1.1 Are there any common in CVD patients?
We choose several factors from the dataset and check if there are some factors that have a significant difference between CVD presented and CVD not presented.All the tables below will use percentage. 

We start with smoke history and whether CVD are currently presented or not.

Smoke
```{r,results='markup'}
library(dplyr)
smoke_con<-aggregate(CVD$cardio, by=list(CVD$smoke,CVD$cardio), FUN=length)
smoke_con<-rename(smoke_con,Smoke = Group.1,CVD_presence = Group.2,Count = x)
smoke_con$Smoke[smoke_con$Smoke=="0"] <- 'Not smoke'
smoke_con$Smoke[smoke_con$Smoke=="1"] <- 'Smoke'
smoke_con$CVD_presence[smoke_con$CVD_presence=="0"] <- 'Not Present'
smoke_con$CVD_presence[smoke_con$CVD_presence=="1"] <- 'Present'
xkabledply(smoke_con)
```

The result showed most respondents do not smoke, and almost half of the respondents have CVD currently presented. 

alcohol intake 
```{r,results='markup'}
alco_con<-aggregate(CVD$cardio,by=list(CVD$alco,CVD$cardio), FUN=length)
alco_con<-rename(alco_con,alco = Group.1,CVD_presence = Group.2,Count = x)
alco_con$alco[alco_con$alco=="0"] <- 'Not drink'
alco_con$alco[alco_con$alco=="1"] <- 'Drink'
alco_con$CVD_presence[alco_con$CVD_presence=="0"] <- 'Not Present'
alco_con$CVD_presence[alco_con$CVD_presence=="1"] <- 'Present'
xkabledply(alco_con)
```

The result showed most respondents do not drink, and almost half of the respondents have CVD currently presented. 

Physical Activity
```{r,results='markup'}
active_con<-aggregate(CVD$cardio, by=list(CVD$active,CVD$cardio), FUN=length)
active_con<-rename(active_con,active = Group.1,CVD_presence = Group.2,Count = x)
active_con$active[active_con$active=="0"] <- 'Physical Activity'
active_con$active[active_con$active=="1"] <- 'Not Physical Activity'
active_con$CVD_presence[active_con$CVD_presence=="0"] <- 'Not Present'
active_con$CVD_presence[active_con$CVD_presence=="1"] <- 'Present'
xkabledply(active_con)
```

The result showed most respondents do not physical activity regularly, and most of the respondents have CVD currently presented. However, there are more people who have physical activity regularly compared to smoking and drinking. 

# 3.1.2 Plot comparing different factors

With all the tables above, we plot graphs accordingly, from the graphs we will have a better idea to see if there are any variables that may have a significant impact on getting Cardio Vascular Diseases. 

Plot of Smoke & CVD
```{r,results='markup'}
smoke_pic <- data.frame(table(CVD$smoke,CVD$cardio))
names(smoke_pic) <- c("Smoke","Cardio","Count")
ggplot(data=smoke_pic, aes(x=Smoke, y=Count/sum(Count)*100, fill=Cardio)) + geom_bar(stat="identity")+scale_x_discrete(labels=c("0" = "Not smoke", "1" = "Smoke"))+scale_fill_discrete(labels=c("0" = "No cvd presented", "1" = "CVD presented"))+ggtitle("Percentage of Smoke history and present of CVDs") + ylab("Percentage")
```

Plot of alcohol intake & CVD
```{r,results='markup'}
alco_pic <- data.frame(table(CVD$alco,CVD$cardio))
names(alco_pic) <- c("alco","Cardio","Count")
ggplot(data=alco_pic, aes(x=alco, y=Count/sum(Count)*100, fill=Cardio)) + geom_bar(stat="identity")+scale_x_discrete(labels=c("0" = "Not Drink", "1" = "Drink"))+scale_fill_discrete(labels=c("0" = "No cvd presented", "1" = "CVD presented"))+ggtitle("Percentage of alcohol intake and present of CVDs") + ylab("Percentage")
```

Plot of physical activity & CVD
```{r,results='markup'}
active_pic <- data.frame(table(CVD$active,CVD$cardio))
names(active_pic) <- c("active","Cardio","Count")
ggplot(data=active_pic, aes(x=active, y=Count/sum(Count)*100, fill=Cardio)) + geom_bar(stat="identity")+scale_x_discrete(labels=c("0" = "Workout regularly", "1" = "Not workout regularly"))+scale_fill_discrete(labels=c("0" = "No cvd presented", "1" = "CVD presented"))+ggtitle("Percentage of workout and present of CVDs") + ylab("Percentage")
```

The graphs proved the results are pretty clear that most respondents do not have a smoking history or drinking history while there are more people who have physical activity regularly. However, from the graphs the chance of getting Cardio-Vascular Diseases seems pretty even regarding their lifestyle. We will use chi-square test below to further test if these variables are independent from whether Cardio-Vascular Diseases are presented. 

# 3.1.3 Chi-Square Tests

The results of the chi-square test result are listed below accordingly. 


Chi-square test for Smoke & CVD presence

Are they independent?
- H0: CVD presence and smoking are independent.
- H1: They are not independent.

```{r,results='hide'}
smoke_cvd = table(CVD$cardio, CVD$smoke)
```

```{r,results='markup'}
sctest<- chisq.test(smoke_cvd)
sctest
```

Since the p-value is 4e-05 < 0.05, we reject the null hypothesis, therefore smoking has a significant impact on whether cardio-vascular diseases are presented. 

Chi-square test for Alcohol intake & CVD presence

Are they independent?
- H0: CVD presence and drinking are independent.
- H1: They are not independent.

```{r,results='hide'}
al_cvd = table(CVD$cardio, CVD$alco)
```
```{r,results='markup'}
alctest<- chisq.test(al_cvd)
alctest
```

Since the p-value is 0.05 ≤ 0.05, we reject the null hypothesis, therefore drink has a significant impact on whether cardio-vascular diseases are presented. 

Chi-square test for Physical Activity & CVD presence

Are they independent?
- H0: CVD presence and physical activity are independent.
- H1: They are not independent.

```{r,results='hide'}
act_cvd = table(CVD$cardio, CVD$active)
```
```{r,results='markup'}
act_test<- chisq.test(act_cvd)
act_test
```

Since the p-value is 2e-16 < 0.05, we reject the null hypothesis, therefore physical activity has a significant impact on whether cardio-vascular diseases are presented. 

# 3.2 Continuous Variables EDA

